Semilattice Graphs for Distributed Databases

### 2022-10-03 paper/paper.tex
we need more parens
circ for readability
T is now L

      \begin{itemize}
        \item For any $x : L, x' : l(x)$: $$t(x, x, x') = x'$$.
        \item For any $x, y, z : L$ such that $x \leq y \leq z$, and $x' : l(x)$: $$t(y, z, t(x, y, x')) = t(x, z, x')$$.
        \item For any $x, y : L$ such that $x \leq y$, and $x'_1, x'_2 : l(x)$: $$t(x, y, x'_1 \vee x'_2) = t(x, y, x'_1) \vee t(x, y, x'_2)$$.
      \end{itemize}

@@

      The total space $\mathrm{SemiTot}(L, l, t)$ is then a semilattice whose elements are pairs $(x, x')$ where $x : L$ and $x' : l(x)$. The join is defined as $(x, x') \vee (y, y') := (x \vee y, t(x, x \vee y, x') \vee t(y, x \vee y, y'))$. This is equivalent to the above category-theoretic definition, proving it is a semilattice.

@@

      Define $T_0$ to be a semilattice with only a single element, $()$. Now suppose we have:

      \begin{align*}
        (l_n, t_n) &: \mathrm{SemiFibration}(T_{n-1}) & \text{ for natural $n \geq 1$ } \\
        T_n &:= \mathrm{SemiTot}(T_{n-1}, l_n, t_n)    & \text{ for natural $n \geq 1$ }
      \end{align*}

      We now wish to define $L_\omega$ to be a semilattice that is, morally speaking, the limit of $L_n$. This lattice's elements are infinite lists $x_1, x_2, \ldots$ satisfying $x_n : l_n(x_{<n})$ for each natural $n \geq 1$, where $x_{<n} = x_1, \ldots, x_{n-1} : T_{n-1}$. The joins are defined as follows:

      $$(\overline{x} \vee \overline{y})_n := t_n(x_{<n}, x_{<n} \vee y_{<n}, x_n) \vee_{l_n(x_{<n} \vee y_{<n})} t_n(y_{<n}, x_{<n} \vee y_{<n}, y_n)$$

@@

        To verify that $L_\omega$ is a semilattice, we must verify three properties:
        \begin{itemize}
          \item \emph{Commutativity}:
            \begin{align*}
              (\overline{x} \vee \overline{y})_n
              &= t_n(x_{<n}, x_{<n} \vee y_{<n})(x_n) \vee_{l_n(x_{<n} \vee y_{<n})} t_n(y_{<n}, x_{<n} \vee y_{<n}, y_n) \\
              &= t_n(y_{<n}, y_{<n} \vee x_{<n})(y_n) \vee_{l_n(y_{<n} \vee x_{<n})} t_n(x_{<n}, y_{<n} \vee x_{<n}, x_n) \\
              &= (\overline{y} \vee \overline{x})_n
            \end{align*}
          \item \emph{Associativity}:
            \begin{align*}
              & ((\overline{x} \vee \overline{y}) \vee \overline{z})_n \\
              =& t_n(x_{<n} \vee y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, (\overline{x} \vee \overline{y})_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})}
                 t_n(z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, z_n)  \\
              =& t_n(x_{<n} \vee y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n},
                          t_n(x_{<n}, x_{<n} \vee y_{<n}, x_n) \vee_{l_n(x_{<n} \vee y_{<n})}
                          t_n(y_{<n}, x_{<n} \vee y_{<n}, y_n)) \\
               &\vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} t_n(z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, z_n)  \\
              =& t_n(x_{<n} \vee y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, t_n(x_{<n}, x_{<n} \vee y_{<n}, x_n)) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
               & t_n(x_{<n} \vee y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, t_n(y_{<n}, x_{<n} \vee y_{<n}, y_n)) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
               & t_n(z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, z_n)  \\
              =& t_n(x_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, x_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
               & t_n(y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, y_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
               & t_n(z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, z_n)  \\
              =& t_n(x_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, x_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
               & t_n(y_{<n} \vee z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, t_n(y_{<n}, y_{<n} \vee z_{<n}, y_n)) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
               & t_n(y_{<n} \vee z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, t_n(z_{<n}, y_{<n} \vee z_{<n}, z_n))  \\
              =& t_n(x_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, x_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
               & t_n(y_{<n} \vee z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n},
                          t_n(y_{<n}, y_{<n} \vee z_{<n}, y_n) \vee_{l_n(y_{<n} \vee z_{<n})}
                          t_n(z_{<n}, y_{<n} \vee z_{<n}, z_n)) \\
              =& t_n(x_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, x_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
               & t_n(y_{<n} \vee z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, (\overline{y} \vee \overline{z})_n) \\
              =& (\overline{x} \vee (\overline{y} \vee \overline{z}))_n
            \end{align*}
          \item \emph{Idempotence}:
            \begin{align*}
              (\overline{x} \vee \overline{x})_n &= t_n(x_{<n}, x_{<n} \vee x_{<n})(x_n) \vee_{l_n(x_{<n} \vee x_{<n}))} t_n(x_{<n}, x_{<n} \vee x_{<n}, x_n) \\
              &= t_n(x_{<n}, x_{<n}, x_n) \\
              &= x_n
            \end{align*}
        \end{itemize}
      \end{proof}

>>>>

      \begin{itemize}
        \item For any $x : L, x' : l(x)$: $$t(x, x) = \mathsf{id}$$.
        \item For any $x, y, z : L$ such that $x \leq y \leq z$, and $x' : l(x)$: $$t(y, z) \circ t(x, y) = t(x, z)$$.
        \item For any $x, y : L$ such that $x \leq y$, and $x'_1, x'_2 : l(x)$: $$t(x, y)(x'_1 \vee x'_2) = t(x, y)(x'_1) \vee t(x, y)(x'_2)$$.
      \end{itemize}

@@

      The total space $\mathrm{SemiTot}(L, l, t)$ is then a semilattice whose elements are pairs $(x, x')$ where $x : L$ and $x' : l(x)$. The join is defined as $(x, x') \vee (y, y') := (x \vee y, t(x, x \vee y)(x') \vee t(y, x \vee y)(y'))$. This is equivalent to the above category-theoretic definition, proving it is a semilattice.

@@

      Let $L_0$ be an arbitrary semilattice. Now suppose we have:

      \begin{align*}
        (l_n, t_n) &: \mathrm{SemiFibration}(L_{n-1}) & \text{ for natural $n \geq 1$ } \\
        L_n &:= \mathrm{SemiTot}(L_{n-1}, l_n, t_n)    & \text{ for natural $n \geq 1$ }
      \end{align*}

      We now wish to define $L_\omega$ to be a semilattice that is, morally speaking, the limit of $L_n$. This lattice's elements are infinite lists $x_1, x_2, \ldots$ satisfying $x_n : l_n(x_{<n})$ for each natural $n \geq 1$, where $x_{<n} = x_1, \ldots, x_{n-1} : L_{n-1}$. The joins are defined as follows:

      $$(\overline{x} \vee \overline{y})_n := t_n(x_{<n}, x_{<n} \vee y_{<n})(x_n) \vee_{l_n(x_{<n} \vee y_{<n})} t_n(y_{<n}, x_{<n} \vee y_{<n})(y_n)$$

@@

        To verify that $L_\omega$ is a semilattice, we must verify three properties:
        \begin{itemize}
          \item \emph{Commutativity}:
            \begin{align*}
              (\overline{x} \vee \overline{y})_n
              &= t_n(x_{<n}, x_{<n} \vee y_{<n})(x_n) \vee_{l_n(x_{<n} \vee y_{<n})} t_n(y_{<n}, x_{<n} \vee y_{<n})(y_n) \\
              &= t_n(y_{<n}, y_{<n} \vee x_{<n})(y_n) \vee_{l_n(y_{<n} \vee x_{<n})} t_n(x_{<n}, y_{<n} \vee x_{<n})(x_n) \\
              &= (\overline{y} \vee \overline{x})_n
            \end{align*}
          \item \emph{Associativity}:
            \begin{align*}
              & ((\overline{x} \vee \overline{y}) \vee \overline{z})_n \\
              =& t_n(x_{<n} \vee y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})((\overline{x} \vee \overline{y})_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})}
                 t_n(z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(z_n)  \\
              =& t_n(x_{<n} \vee y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(
                          t_n(x_{<n}, x_{<n} \vee y_{<n})(x_n) \vee_{l_n(x_{<n} \vee y_{<n})}
                          t_n(y_{<n}, x_{<n} \vee y_{<n})(y_n)) \\
               &\vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} t_n(z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(z_n)  \\
              =& (t_n(x_{<n} \vee y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}) \circ t_n(x_{<n}, x_{<n} \vee y_{<n}))(x_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
               & (t_n(x_{<n} \vee y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}) \circ t_n(y_{<n}, x_{<n} \vee y_{<n}))(y_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
               & t_n(z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(z_n)  \\
              =& t_n(x_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(x_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
               & t_n(y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(y_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
               & t_n(z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(z_n)  \\
              =& t_n(x_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(x_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
               & (t_n(y_{<n} \vee z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}) \circ t_n(y_{<n}, y_{<n} \vee z_{<n}))(y_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
               & (t_n(y_{<n} \vee z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}) \circ t_n(z_{<n}, y_{<n} \vee z_{<n}))(z_n)  \\
              =& t_n(x_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(x_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
               & t_n(y_{<n} \vee z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(
                          t_n(y_{<n}, y_{<n} \vee z_{<n})(y_n) \vee_{l_n(y_{<n} \vee z_{<n})}
                          t_n(z_{<n}, y_{<n} \vee z_{<n})(z_n)) \\
              =& t_n(x_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(x_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
               & t_n(y_{<n} \vee z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})((\overline{y} \vee \overline{z})_n) \\
              =& (\overline{x} \vee (\overline{y} \vee \overline{z}))_n
            \end{align*}
          \item \emph{Idempotence}:
            \begin{align*}
              (\overline{x} \vee \overline{x})_n &= t_n(x_{<n}, x_{<n} \vee x_{<n})(x_n) \vee_{l_n(x_{<n} \vee x_{<n}))} t_n(x_{<n}, x_{<n} \vee x_{<n})(x_n) \\
              &= t_n(x_{<n}, x_{<n})(x_n) \\
              &= x_n
            \end{align*}
        \end{itemize}
      \end{proof}
